---
title: "all sim"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("MASS",character.only=TRUE)
library("ggplot2")
set.seed(10)
```

First we define the needed parameters for the synthetic data. Note that each vector has only two elements.

```{r}
beta<-c(1,2)
rho<-c(0.98,0.95)
a<-0
for (i in 1:1){
  a<-a+rho[i]^2
}
a^(1/2)

sigma<-1
nd<-length(beta)
```

Now we define a function that ensure that the synthetic data is non-stationary.

```{r}
betaupd<-function(rho,beta,sigma){
  u<-rnorm(2,mean=0,sd=sigma)
  newbeta<-(rho*beta) + u
  return(newbeta)
}

```

We now create the synthetic data.

```{r}
num<-100000
d<-length(beta)
x<-mvrnorm(num,rep(1,d),diag(d))
ysim<-function(num,beta,x){
  dim<-length(beta)
  eps<-rnorm(n=num, mean=0, sd=1)
  y<-eps
  for (j in 1:num){
    for (i in 1:dim){
      y[j]<-y[j]+beta[i]*x[j,i]
    }
    beta<-betaupd(rho,beta,sigma)
  }
  return(y)
}
y<-ysim(num,beta,x)
x2<-x
x<-cbind(x,matrix(rep(1,num),ncol=1))
```

Here we code and run the Covariance Method.

```{r}

###########################
# My version
###########################
siginffunc<-function(M, tt){
    M2<-matrix(M,nrow=2,ncol=2)
    tot<-0
    for (i in 1:(tt-1)){
      xi<-x2[i,]
      tot<-tot+((y[i]^2)-(t(xi)%*%M2%*%xi)-(sigeps^2))^2
    }
    return(tot)
  }
  
  
Mfunc<-function(M, tt){
    M2<-matrix(M,nrow=2,ncol=2)
    tot<-0
    for (i in 1:(tt-1)){
      xi<-x2[i,]
      tot<-tot+((y[i]*y[i+1])-(t(x2[i,])%*%M2%*%x2[(i+1),]))^2
    }
    return(tot)
}
ahtfind<-function(t_val,M0, sigma,x,y){
   M<-matrix(optim(t(c(M0)),Mfunc, tt=t_val,   control=list(maxit=100000))$par,nrow=2,ncol=2)
   
 
   siginfhat<-matrix(optim(t(c(M)),siginffunc,tt=t_val, control=list(maxit=100000))$par,nrow=2,ncol=2)
   return(M%*%solve(siginfhat))
}

sigeps=sigma
M0<-matrix(diag(2),2,2)

ahat<-ahtfind(10000,M0, sigma,x2,y)
ahat

 

```
 
Here we code the kalman filter which will allow us to make predictions with our estimate ahat found using CM.

```{r}

kalmanfilt<-function(A,Q,R,x,y){
  mu<-matrix(rep(0,2),nrow=2,ncol=1)
  tt<-length(y)
  lpmu<-mu
  store<-numeric(tt)
  cov<-Q
  for(i in 1:(tt-1)){
    H<-matrix(x[i,],nrow=1)
    pmu<-A%*%mu              #ok x_{k|k-1}
    store[i]<-H%*%pmu
    pcov<-(A%*%cov%*%t(A))+Q #P_{t+1|t}
    K<-(pcov%*%t(H))%*%(solve(H%*%pcov%*%t(H)+R))
    mu<-pmu+K*c(y[i]-H%*%pmu) 
    cov<-pcov- K%*%H%*%pcov
  }
  return(store)
}

A0<-diag(rho)

yhatkal<-kalmanfilt(A=ahat, Q=diag(rep(sigma,2)), R=sigma,x2,y)
 
```

We now build a function that will calculate the cumulative squared loss.

```{r}
cumsqloss<-function(y,yh,burn){
  Tt<-length(y)
  return(sum((y[(burn+1):Tt]-yh[(burn+1):Tt])^2))
}

cumsqloss2<-function(y,yh){
  Tt<-length(y)
  tot<-0
  store<-numeric(Tt)
  for (i in (1):Tt){
    tot<-tot+(yh[i]-y[i])^2
    store[i]<-tot
  }
  return(store)
}
```

We now run the Kalman filter with our estimate ahat found using the CM on the full data set and analyse them accordingly.

```{r}
yhatkal2<-yhatkal[10001:num]

ytest<-y[10001:num]

cumsqloss(ytest,yhatkal2,0)/(num-10000)

plot((1:(num-10000)),(ytest-yhatkal2))

cumkal<-cumsqloss2(ytest,yhatkal2)

plot((1:(num-10000)),cumkal,type="l",col="red")
```

We now code all the other algorithms previously discussed so we can compare.

```{r}
alphast<-function(u,alpha,lamda,Rb){
  n<-length(alpha)
  store<-numeric(n)
  for (i in 1:n){
    store[i]<-sum((u^2)/((1+alpha[i]*lamda)^2))
  }
  return(store)
}

proj<-function(wt,covv,Rb,dim){
  decomp<-eigen(covv)
  V<-decomp$vectors
  Lamd<-decomp$values
  u<-t(V)%*%wt
  if (norm(wt,type="2")<=Rb){
    return(wt)
  }
  else{
    lamdmin<-min(Lamd)
    a<-((norm(u,type="2"))/(Rb-1))/lamdmin
    alpha<-seq(0,a,a/1000)
    alpha<-alpha[binary_search(alphast(u,alpha,Lamd,Rb),0,TRUE)]
    return(solve((diag(dim))+alpha*covv)%*%wt)
  }
  return(w)
}

ARCOR<-function(x,y,r,Rb,ll){
  dim<-length(x[1,])
  num<-length(x[,1])
  I<-diag(dim)
  tcov<-ll*I
  w<-t(matrix(0,1,dim))
  covv<-diag(dim)
  ey<-c()
  for (i in 1:num){
    denom<-r+t(x[i,])%*%covv%*%x[i,]
    numer<-t(x[i,])%*%w
    ey[i]<-numer
    wnum<-(y[i]-numer)
    wtilde<-w+(wnum/denom)[1,1]*covv%*%x[i,]
    covvtilde<-solve(solve(covv)+(x[i,]%*%t(x[i,])))
    if (min((as.vector(covv)-as.vector(tcov)))>=0){
      covv<-covvtilde
    }
    else {
      covv<-I
      tcov<-(ll*(4/5))*I
    }
    w<-proj(wtilde,covv,Rb,dim)
  }
  return(ey)
}

AROWR<-function(x,y,r){
  dim<-length(x[1,])
  num<-length(x[,1])
  w<-t(matrix(0,1,dim))
  covv<-diag(dim)
  ey<-c()
  for (i in 1:num){
    ey[i]<-t(x[i,])%*%w
    wnum<-(y[i]-ey[i])
    wden<-r+t(x[i,])%*%covv%*%x[i,]
    w<-w+(wnum/wden)[1,1]*covv%*%x[i,]
    covv<-solve(solve(covv)+(1/r)*(x[i,]%*%t(x[i,])))
  }
  return(ey)
}
```


```{r}
LASER<-function(x,y,b,c){
  dim<-length(x[1,])
  num<-length(x[,1])
  w<-t(matrix(0,1,dim))
  I<-diag(dim)
  covv<-I*((c-b)/(b*c))
  ey<-c()
  for (i in 1:num){
    denom<-1+t(x[i,])%*%(covv+((1/c)*I))%*%x[i,]
    numer<-t(x[i,])%*%w
    ey[i]<-numer/denom
    wnum<-(y[i]-numer)
    w<-w+(wnum/denom)[1,1]*(covv+((1/c)*I))%*%x[i,]
    covv<-solve(solve((covv+((1/c)*I)))+(x[i,]%*%t(x[i,])))
  }
  return(ey)
}

AAR<-function(x,y,b){
  dim<-length(x[1,])
  num<-length(x[,1])
  w<-t(matrix(0,1,dim))
  covv<-diag(dim)*(1/b)
  ey<-c()
  for (i in 1:num){
    denom<-1+t(x[i,])%*%covv%*%x[i,]
    numer<-t(x[i,])%*%w
    ey[i]<-numer/denom
    wnum<-(y[i]-numer)
    w<-w+(wnum/denom)[1,1]*covv%*%x[i,]
    covv<-solve(solve(covv)+(x[i,]%*%t(x[i,])))
  }
  return(ey)
}
```


```{r}
CRRLS<-function(x,y,r,t0){
  dim<-length(x[1,])
  num<-length(x[,1])
  w<-t(matrix(0,1,dim))
  covv<-diag(dim)
  ey<-c()
  for (i in 1:num){
    ey[i]<-t(x[i,])%*%w
    wnum<-(y[i]-ey[i])
    wden<-r+t(x[i,])%*%covv%*%x[i,]
    w<-w+(wnum/wden)[1,1]*covv%*%x[i,]
    covv<-solve(r*solve(covv)+(x[i,]%*%t(x[i,])))
    if (i%%t0==0){
      covv<-diag(dim)
    }
  }
  return(ey)
}

RLS<-function(x,y,r){
  dim<-length(x[1,])
  num<-length(x[,1])
  w<-t(matrix(0,1,dim))
  covv<-diag(dim)
  ey<-c()
  for (i in 1:num){
    ey[i]<-t(x[i,])%*%w
    wnum<-(y[i]-ey[i])
    wden<-r+t(x[i,])%*%covv%*%x[i,]
    w<-w+(wnum/wden)[1,1]*covv%*%x[i,]
    covv<-solve(r*solve(covv)+(x[i,]%*%t(x[i,])))
  }
  return(ey)
}
```

We now find the optimum parameters using a training set (this time of the first 10,000 data entries) and then run the algorithms with these parameters and perform the corresponding analysis. We do this for all algorithms.

```{r}
cumsqloss(y[500:10000],LASER(x,y,1,100)[500:10000],0)
cumsqloss(y[500:10000],LASER(x,y,1,200)[500:10000],0)
cumsqloss(y[500:10000],LASER(x,y,1,300)[500:10000],0)
cumsqloss(y[500:10000],LASER(x,y,1,400)[500:10000],0)
cumsqloss(y[500:10000],LASER(x,y,1,500)[500:10000],0)
cumsqloss(y[500:10000],LASER(x,y,1,600)[500:10000],0)
cumsqloss(y[500:10000],LASER(x,y,1,700)[500:10000],0)
cumsqloss(y[500:10000],LASER(x,y,1,800)[500:10000],0)
cumsqloss(y[500:10000],LASER(x,y,1,900)[500:10000],0)
cumsqloss(y[500:10000],LASER(x,y,1,1000)[500:10000],0)
```

```{r}
b<-1
c<-200
b2<-(b*c)/(c-b)

yhatLASER<-LASER(x,y,b,c)[10001:num]
yhatAAR<-AAR(x,y,b2)[10001:num]

ytest<-y[10001:num]

cumsqloss(ytest,yhatLASER,0)/(num-10000)
cumsqloss(ytest,yhatAAR,0)/(num-10000)

plot((1:(num-10000)),(ytest-yhatAAR))
plot((1:(num-10000)),(ytest-yhatLASER))

cumAAR<-cumsqloss2(ytest,yhatAAR)
cumLASER<-cumsqloss2(ytest,yhatLASER)
plot((1:(num-10000)),cumAAR,type="l",col="red")
lines((1:(num-10000)),cumLASER,col="green")
```

```{r}
cumsqloss(y[500:10000],ARCOR(x,y,1,10000,0.95)[500:10000],0)
cumsqloss(y[500:10000],ARCOR(x,y,2,10000,0.95)[500:10000],0)
cumsqloss(y[500:10000],ARCOR(x,y,3,10000,0.95)[500:10000],0)
cumsqloss(y[500:10000],ARCOR(x,y,4,10000,0.95)[500:10000],0)
cumsqloss(y[500:10000],ARCOR(x,y,5,10000,0.95)[500:10000],0)
cumsqloss(y[500:10000],ARCOR(x,y,6,10000,0.95)[500:10000],0)
cumsqloss(y[500:10000],ARCOR(x,y,100,10000,0.95)[500:10000],0)
```

```{r}
cumsqloss(y[500:10000],AROWR(x,y,1)[500:10000],0)
cumsqloss(y[500:10000],AROWR(x,y,1000)[500:10000],0)
cumsqloss(y[500:10000],AROWR(x,y,2000)[500:10000],0)
cumsqloss(y[500:10000],AROWR(x,y,3000)[500:10000],0)
cumsqloss(y[500:10000],AROWR(x,y,4000)[500:10000],0)
cumsqloss(y[500:10000],AROWR(x,y,5000)[500:10000],0)
cumsqloss(y[500:10000],AROWR(x,y,6000)[500:10000],0)
cumsqloss(y[500:10000],AROWR(x,y,7000)[500:10000],0)
cumsqloss(y[500:10000],AROWR(x,y,8000)[500:10000],0)
cumsqloss(y[500:10000],AROWR(x,y,9000)[500:10000],0)
cumsqloss(y[500:10000],AROWR(x,y,10000)[500:10000],0)
```

```{r}
yhatARCOR<-ARCOR(x,y,1,100,0.95)[10001:num]
yhatAROWR<-AROWR(x,y,4000)[10001:num]

ytest<-y[10001:num]

cumsqloss(ytest,yhatARCOR,0)/(num-10000)
cumsqloss(ytest,yhatAROWR,0)/(num-10000)

plot((1:(num-10000)),(ytest-yhatAROWR))
plot((1:(num-10000)),(ytest-yhatARCOR))

cumAROWR<-cumsqloss2(ytest,yhatAROWR)
cumARCOR<-cumsqloss2(ytest,yhatARCOR)
plot((1:(num-10000)),cumAROWR,type="l",col="red")
lines((1:(num-10000)),cumARCOR,col="green")
```
```{r}
cumsqloss(y[500:10000],CRRLS(x,y,0.9,1)[500:10000],0)
cumsqloss(y[500:10000],CRRLS(x,y,0.9,2)[500:10000],0)
cumsqloss(y[500:10000],CRRLS(x,y,0.9,3)[500:10000],0)
cumsqloss(y[500:10000],CRRLS(x,y,0.9,4)[500:10000],0)
cumsqloss(y[500:10000],CRRLS(x,y,0.9,5)[500:10000],0)
cumsqloss(y[500:10000],CRRLS(x,y,0.9,10)[500:10000],0)
cumsqloss(y[500:10000],CRRLS(x,y,0.9,15)[500:10000],0)
cumsqloss(y[500:10000],CRRLS(x,y,0.9,20)[500:10000],0)
cumsqloss(y[500:10000],CRRLS(x,y,0.9,25)[500:10000],0)
cumsqloss(y[500:10000],CRRLS(x,y,0.9,30)[500:10000],0)
cumsqloss(y[500:10000],CRRLS(x,y,0.9,35)[500:10000],0)
cumsqloss(y[500:10000],CRRLS(x,y,0.9,40)[500:10000],0)
cumsqloss(y[500:10000],CRRLS(x,y,0.9,45)[500:10000],0)
cumsqloss(y[500:10000],CRRLS(x,y,0.9,50)[500:10000],0)
cumsqloss(y[500:10000],CRRLS(x,y,0.9,10000)[500:10000],0)
```

```{r}
yhatCRRLS<-CRRLS(x,y,0.9,2)[10001:num]
yhatRLS<-RLS(x,y,0.9)[10001:num]

ytest<-y[10001:num]

cumsqloss(ytest,yhatCRRLS,0)/(num-10000)
cumsqloss(ytest,yhatRLS,0)/(num-10000)

plot((1:(num-10000)),(ytest-yhatRLS))
plot((1:(num-10000)),(ytest-yhatCRRLS))

cumRLS<-cumsqloss2(ytest,yhatRLS)
cumCRRLS<-cumsqloss2(ytest,yhatCRRLS)
plot((1:(num-10000)),cumRLS,type="l",col="red")
lines((1:(num-10000)),cumCRRLS,col="green")
```

We now plot all cumulative squared errors on one glass for easy comparison

```{r}
plot((1:(num-10000)),cumRLS,type="l",col="red",main="Synthetic Data",ylab="Cumulative Squared-Error",xlab="Number of iterations")
lines((1:(num-10000)),cumCRRLS,col="orange")
lines((1:(num-10000)),cumARCOR,col="purple")
lines((1:(num-10000)),cumAROWR,col="blue")
lines((1:(num-10000)),cumAAR,col="darkgreen")
lines((1:(num-10000)),cumLASER,col="green")
lines((1:(num-10000)),cumkal,col="pink")
legend(0, 1800000, legend=c("RLS", "CR-RLS","ARCOR", "AROWR","AAR", "LASER","CM + Kalman Filter"),
       col=c("red", "orange","purple", "blue","darkgreen", "green","pink"), lty=1:2, cex=0.8)
```



