---
title: "all sim"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("MASS",character.only=TRUE)
library("ggplot2")
set.seed(10)
```

First we define the needed parameters for the synthetic data. Note that each vector has only two elements.

```{r}
beta<-c(1,3,2,4,5,3,1,3,2,1)
rho<-c(0.98,0.95,0.97,0.99,0.95,1,0.98,0.99,0.95,0.96)
a<-0
for (i in 1:1){
  a<-a+rho[i]^2
}
a^(1/2)

sigma<-1
nd<-length(beta)
```

Now we define a function that ensure that the synthetic data is non-stationary.

```{r}
betaupd<-function(rho,beta,sigma){
  u<-rnorm(nd,mean=0,sd=sigma)
  newbeta<-(rho*beta) + u
  return(newbeta)
}

betaupd2<-function(c,t){
  newbeta2<-beta*sin(c*t)
  return(newbeta2)
}
```

We now create the synthetic data.

```{r}
num<-100000
d<-length(beta)
x<-mvrnorm(num,rep(1,d),diag(d))
ysim<-function(num,beta,x){
  dim<-length(beta)
  eps<-rnorm(n=num, mean=0, sd=1)
  y<-eps
  for (j in 1:num){
    for (i in 1:dim){
      y[j]<-y[j]+beta[i]*x[j,i]
    }
    beta<-betaupd(rho,beta,sigma)
  }
  return(y)
}
y<-ysim(num,beta,x)
x2<-x
x<-cbind(x,matrix(rep(1,num),ncol=1))
```

Here we code and run the Covariance Method.

```{r}

###########################
# My version
###########################
siginffunc<-function(M, tt){
    M2<-diag(M)
    tot<-0
    for (i in 1:(tt-1)){
      xi<-x2[i,]
      tot<-tot+((y[i]^2)-(t(xi)%*%M2%*%xi)-(sigeps^2))^2
    }
    return(tot)
  }
  
  
Mfunc<-function(M, tt){
    M2<-diag(M)
    tot<-0
    for (i in 1:(tt-1)){
      xi<-x2[i,]
      tot<-tot+((y[i]*y[i+1])-(t(x2[i,])%*%M2%*%x2[(i+1),]))^2
    }
    return(tot)
}
ahtfind<-function(t_val,M0, sigma,x,y){
   M<-diag(optim(M0,Mfunc, tt=t_val,   control=list(maxit=1000000))$par)
   
 
   siginfhat<-diag(optim(M0,siginffunc,tt=t_val, control=list(maxit=1000000))$par)
   return(M%*%solve(siginfhat))
}

sigeps=sigma
M0<-rep(1,nd)

ahat<-ahtfind(100000,M0, sigma,x2,y)
ahat
```
 
Here we code the kalman filter which will allow us to make predictions with our estimate ahat found using CM.

```{r}

kalmanfilt<-function(A,Q,R,x,y){
  mu<-matrix(rep(0,nd),nrow=nd,ncol=1)
  tt<-length(y)
  lpmu<-mu
  store<-numeric(tt)
  cov<-Q
  for(i in 1:(tt-1)){
    H<-matrix(x[i,],nrow=1)
    pmu<-A%*%mu              #ok x_{k|k-1}
    store[i]<-H%*%pmu
    pcov<-(A%*%cov%*%t(A))+Q #P_{t+1|t}
    K<-(pcov%*%t(H))%*%(solve(H%*%pcov%*%t(H)+R))
    mu<-pmu+K*c(y[i]-H%*%pmu) 
    cov<-pcov- K%*%H%*%pcov
  }
  return(store)
}

A0<-diag(rho)

yhatkal<-kalmanfilt(A=ahat, Q=diag(rep(sigma,nd)), R=sigma,x2,y)
 
```

We now build a function that will calculate the cumulative squared loss.

```{r}
cumsqloss<-function(y,yh,burn){
  Tt<-length(y)
  return(sum((y[(burn+1):Tt]-yh[(burn+1):Tt])^2))
}

cumsqloss2<-function(y,yh){
  Tt<-length(y)
  tot<-0
  store<-numeric(Tt)
  for (i in (1):Tt){
    tot<-tot+(yh[i]-y[i])^2
    store[i]<-tot
  }
  return(store)
}
```

We now run the Kalman filter with our estimate ahat found using the CM on the full data set and analyse them accordingly.

```{r}
yhatkal2<-yhatkal[10001:num]

ytest<-y[10001:num]

cumsqloss(ytest,yhatkal2,0)/(num-10000)

plot((1:(num-10000)),(ytest-yhatkal2))

cumkal<-cumsqloss2(ytest,yhatkal2)

plot((1:(num-10000)),cumkal,type="l",col="red")
```



